---
title: "Exploratory Analysis"
output: html_document
---

## Project title

In the Dog Pound

### Authors

Olivia Do, Cathy Nguyen, Jenny Suk, and Leo Zheng

### Date

Winter 2022

### Abstract

Our focus is on the population of dogs in shelters. Focusing on the statistics of certain breed of dogs in the shelter, the gender, age, coat and whether or not they are fixed, to see the duration of how long they tend to stay in the shelter. To address this subject we will compare the statistics provided from the alldogdescprition data provided from the pudding dataset with each other.

### Keywords

Dogs Shelter Statistics

### Introduction  

Our main focus is the statistics behind dogs in shelters; as a group, we want to understand the impact that certain variables have on the chances for a dog to get adopted. We want to see how the age of the dogs affects the chances of it getting a family. Whether the dog is female or male impacts the likelihood of adoption. The breed, to when they were posted for adoption, as well as the coat of the dog having an effect on getting chosen. We will be extracting the data from alldogdescription, in able to see the statistics regarding how our given variables will affect the chances of adoption. We were motivated to look at the population of dogs in shelters to see how the characteristics of the dog can impact their ability to get adopted and what actions that we are able to take to increase the likelihood of these dogs to get a better life. This is important because dogs have the ability to improve the lives of people and by providing these dogs with a home they are able to change their lives as well. Our goal is to present the information for dog owners and future dog owners to be mindful of when providing our furry friends a chance to change their lives.

### Related Work  

The first decision to be made is the place of adoption, whether that is from a breeder or shelter. In the shelter, there is variety in dog breeds to select from, enabling future dog owners to see who the best fit for them would be. Since each dog has unique characteristics and personalities as well as each dog-owner having certain requirements when adopting a pet, we aim to investigate how these differences affect their appeal or lack thereof. Some dog breeds suffer from negative reputations and bad connotations regarding their behaviors caused by these misconstrued stereotypes. Common misconceptions among non-dog owners label rottweilers as aggressive and rough when in reality they are large and excitable. Another example comes from chihuahuas, which are perceived to be annoying and noisy when they just lack proper training. Understandably, people could have allergies that limit their options and decision in the search for a canine companion. Dogs with curly and longer coats tend to shed less and are hypoallergenic, although that does not mean they are completely allergy-free, they do significantly reduce the symptoms and reactions to fur. Lastly, the time a potential dog owner is willing to commit makes up argumentatively the most important step in adoption. Raising a dog is a big responsibility, some lead busier lives than others, have less time to dedicate to training and prefer low maintenance breeds meanwhile a family with kids could keep up with high energy ones that require a lot of attention. All these factors need to be taken in consideration when evaluating why a specific breed is adopted more or less than another should a correlation be found.

References

Editor. “Debunking 5 Common Dog Breed Stereotypes.” Dog's Best Life, DogsBestLife.com Online Magazine, 8 Feb. 2022, https://dogsbestlife.com/dog-breeds/dog-breed-stereotypes/.

Schamble, Melody. “Dog Breeds That Don't Shed: Best Dogs for Allergies.” American Kennel Club, American Kennel Club, 1 July 2022, https://www.akc.org/expert-advice/lifestyle/dog-breeds-that-dont-shed/.

Stregowski, Jenna. “10 Low-Maintenance Dog Breeds for Busy or Lazy Owners.” The Spruce Pets, The Spruce Pets, 21 Sept. 2022, https://www.thesprucepets.com/low-maintenance-dog-breeds-4145311.

### The Dataset

Where did you find the data? Please include a link to the data source  
Who collected the data?  
How was the data collected or generated?  
Why was the data collected?  
How many observations (rows) are in your data?  
How many features (columns) are in the data?  
What, if any, ethical questions or questions of power do you need to consider when working with this data?  
What are possible limitations or problems with this data?   (at least 200 words)

### Implications

Assuming you answer your research questions, briefly describe the expected or possible implications for technologists, designers, and policymakers. (at least 150 words)

### Limitations & Challenges
What challenges or limitations might you need to address with your project idea more broadly? Briefly discuss. (at least 150 words)

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

### Table

```{r, echo=FALSE}
newdog <- allDogDescriptions %>% 
  select()
```

Describe why you included the table and what information it reveals

### Chart 1

Include a chart

library(ggplot2)

Describe why you chose this chart and what information it reveals

Here's an example of how to run an R script inside an RMarkdown file:
Problems: getting the color to go on the bars, and changing the order of the bars
```{r, echo=FALSE}
ggplot(allDogDescriptions, aes(x = age, fill = age)) +
  geom_bar()+
  labs(title = "Age of Dogs in Shelters",
       x = "Age",
       y = "Quantity of Dogs in Shelters") 
```


In order to visually let our audience see the categorical data difference between the age range of the dogs in the shelter, we decided that a bar chart was the best method. By using the bar chart we are able to compare the stages of life that each dog is at through categorizing it by; baby, young, adult, and senior. We also added color to each bar to allow viewers a better visual comparison. The information in this chart reveals a correlation between the age of dogs and how many get left in the shelter. As a dog gets older, we see a larger quantity left in the shelter, however, there is the outlier of senior dogs that we see have the lowest quantity in the shelter. Alternatively, there are other factors that can contribute to this, for example dying of old age.

### Chart 2

Include a chart

Describe why you chose this chart and what information it reveals

### Chart 3

Include a chart
```{r chart 3, echo=FALSE}
source('histogram.R')
hist

```

Describe why you chose this chart and what information it reveals

By splitting this histogram into different facets, we attempt to show the number of days that a dog typically remains in a shelter based on their gender Using a histogram helps visualize the frequency distribution of the dataset making it is easier to tell how many days are the most or least common for a dog to stay for. The graphs reveal that a majority of dogs stay for around 1200 days. Additionally, they are skewed right, meaning that the number of dogs that stay past 1200 days decreases as time goes on. Several possible outliers exist as well, which are most likely those that have stayed for over 6000 days. Another piece of information that this visualization expresses is that a dog's gender does not play a large role in its adoption since we see that the mode heavily outnumbers the other values. 
